---
title: "Zooplankton Microbiomes Across Time and Space"
author: "MG Perreault, CB Wall"
date: "1/15/2024"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
editor_options: 
  chunk_output_type: console
---
```{r global options, results="hide", warning=FALSE, message=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

devtools::install_github("benjjneb/dada2", ref="v1.20") # update to most recent dada2
devtools::install_github("zdk123/SpiecEasi")
remotes::install_github("microbiome/microbiome")


# use pacman to load CRAN packages missing
pacman::p_load('knitr', 'microbiome', 'phyloseq', 'tidyr', 'tidyverse', 'knitr', 'magrittr', 'effects', 'devtools',
               'stringi', 'dplyr', "ggplot2", "gridExtra", "dada2", "phyloseq", "vegan", "cowplot", 'doBy', 'ecodist',
               'glue', 'geosphere', 'data.table', 'patchwork', 'car', 'ggcorrplot', 'FactoMineR', 'devtools', 'reshape',
               'lattice',  'plyr', 'magrittr', 'factoextra', 'multcompView', 'decontam', 'factoextra', 'car', 'mia', 
               'ade4', 'fossil', 'picante', 'reshape', 'readr', 'corrr', 'scipplot', 'Hmisc', "MASS", "FSA", "sciplot",
               "decontam","BiocManager", 'ggpubr', 'ggmap', "ggordiplots", "fossil", "SpiecEasi", "igraph", "huge",
               "leaflet")
```

## Overview

### Import Phyloseq
Import the data and modify the metadata to meet downstream needs
```{r loading in data and modifying metadata}
##########################################################

# ps.prune is the final phyloseq object from pipeline, can load it in here...
### or load in the raw data and re-assemble

ps.prune.LOCAL<-readRDS("output/reanalysis/ps.prune_local.RDS")

# idiot check: make sure to remove all mitochondria and chloroplast taxonomic IDs
ps.prune.LOCAL <- subset_taxa(ps.prune.LOCAL, Family!= "Mitochondria" | 
                        is.na(Family) & Class!="Chloroplast" | is.na(Class))

# order the lakes by North to South,  in phyloseq 
sample_data(ps.prune.LOCAL)$Lake <- factor(sample_data(ps.prune.LOCAL)$Lake, levels = 
                                           c("Cooney", "Blue", "Virginia", "Convict", "Serene", "Eastern Brook"))

# export metadata and play
metaD<-microbiome::meta(ps.prune.LOCAL)

# adjust caps
metaD<-metaD %>% 
  dplyr::rename(
    sequencing_ID = Sequencing_ID,
    sample_type = Sample_Type,
    organism = Organism,
    time_point= Time.Point,
    lake = Lake,
    sample_ID = Original_ID
  )

# correct NAs to be "water"
metaD$organism<-as.character(metaD$organism) # need to do this to change the level
metaD$organism[is.na(metaD$organism)] = "Water"
metaD$organism<-as.factor(metaD$organism)
metaD$organism<-factor(metaD$organism, levels =c("Bosmina", "Ceriodaphnia", "Daphnia", "Holopedium", 
                                             "Cyclopoid", "Calanoid", "Large Calanoid",
                                             "Water")) # back to factor


# format metadata (have to redo this if loading back in)
make.fac<-c("sample_ID", "sequencing_ID", "sample_type", "time_point", "lake", "Plate", "Plate_name", "Well", "sample_control")

metaD[make.fac]<-lapply(metaD[make.fac], factor) # make all these factors

# create phy group (Cladocera vs Copepoda vs water) column
metaD$phy_group <- as.factor(ifelse(metaD$organism == "Calanoid" | 
                            metaD$organism == "Cyclopoid" | 
                            metaD$organism == "Large Calanoid", "copepoda", 
                   ifelse(metaD$organism == "Daphnia" | 
                            metaD$organism == "Ceriodaphnia" | 
                            metaD$organism =="Bosmina" | 
                            metaD$organism == "Holopedium", "cladocera", 
                            "water")))

# create basin column
metaD$basin <- as.factor(ifelse(metaD$lake == "Eastern Brook" | 
                                metaD$lake == "Serene" | 
                                metaD$lake == "Convict", "South", 
                                "North"))

# create latitude column
metaD$latitude <- as.numeric(ifelse(metaD$lake == "Blue", "38.050952", 
                            (ifelse(metaD$lake == "Convict", "37.590094",
                            (ifelse(metaD$lake == "Cooney", "38.04806",
                            (ifelse(metaD$lake == "Eastern Brook", "37.431526", 
                            (ifelse(metaD$lake == "Serene", "37.438392" ,
                            (ifelse(metaD$lake == "Virginia", "38.046975",
                                  "0"))))))))))))

# create longitude column
metaD$longitude <- as.numeric(ifelse(metaD$lake == "Blue", "-119.270331", 
                            (ifelse(metaD$lake == "Convict", "-118.857422",
                            (ifelse(metaD$lake == "Cooney", "-119.27702",
                            (ifelse(metaD$lake == "Eastern Brook", "-118.742615", 
                            (ifelse(metaD$lake == "Serene", "-118.744386" ,
                            (ifelse(metaD$lake == "Virginia", "-119.265076",
                                  "0"))))))))))))
                            


##### Read counts and rarefaction curves
# Make a data frame with a column for the read counts of each sample
sample_sum_df<-as.data.frame(sample_sums(ps.prune.LOCAL))
colnames(sample_sum_df)<-"read.sum"
sample_sum_df$sampleNames<-rownames(sample_sum_df)

# are row names the same? if so, add them in
identical(rownames(sample_sum_df), rownames(metaD))
metaD$read.sum<-sample_sum_df$read.sum

# add in dates, as per metadata
metaD$date<-ifelse(metaD$time_point =="1" | metaD$lake =="Blue", "6/29/2022",
                            ifelse(metaD$time_point =="2" | metaD$lake =="Blue", "7/09/2022",
                            ifelse(metaD$time_point =="3" | metaD$lake =="Blue", "7/21/2022",
                            ifelse(metaD$time_point =="4" | metaD$lake =="Blue", "8/04/2022",
                            ifelse(metaD$time_point =="5" | metaD$lake =="Blue", "8/22/2022",
                    
                     ifelse(metaD$time_point =="1" | metaD$lake =="Convict", "6/25/2022",
                            ifelse(metaD$time_point =="2" | metaD$lake =="Convict", "7/07/2022",
                            ifelse(metaD$time_point =="3" | metaD$lake =="Convict", "7/22/2022",
                            ifelse(metaD$time_point =="4" | metaD$lake =="Convict", "8/05/2022",
                            ifelse(metaD$time_point =="5" | metaD$lake =="Convict", "8/23/2022",
                          
                     ifelse(metaD$time_point =="1" | metaD$lake =="Cooney", "6/29/2022",
                            ifelse(metaD$time_point =="2" | metaD$lake =="Cooney", "7/09/2022",
                            ifelse(metaD$time_point =="3" | metaD$lake =="Cooney", "7/21/2022",
                            ifelse(metaD$time_point =="4" | metaD$lake =="Cooney", "8/04/2022",
                            ifelse(metaD$time_point =="5" | metaD$lake =="Cooney", "8/22/2022", 
                                
                     ifelse(metaD$time_point =="1" | metaD$lake =="Eastern Brook", "6/26/2022",
                            ifelse(metaD$time_point =="2" | metaD$lake =="Eastern Brook", "7/06/2022",
                            ifelse(metaD$time_point =="3" | metaD$lake =="Eastern Brook", "7/22/2022",
                            ifelse(metaD$time_point =="4" | metaD$lake =="Eastern Brook", "8/09/2022",
                            ifelse(metaD$time_point =="5" | metaD$lake =="Eastern Brook", "8/23/2022",
                                   
                    ifelse(metaD$time_point =="1" | metaD$lake =="Serene", "6/26/2022",
                            ifelse(metaD$time_point =="2" | metaD$lake =="Serene", "7/06/2022",
                            ifelse(metaD$time_point =="3" | metaD$lake =="Serene", "7/22/2022",
                            ifelse(metaD$time_point =="4" | metaD$lake =="Serene", "8/09/2022",
                            ifelse(metaD$time_point =="5" | metaD$lake =="Serene", "8/23/2022",
                                  
                    ifelse(metaD$time_point =="1" | metaD$lake =="Virginia", "6/29/2022",
                            ifelse(metaD$time_point =="2" | metaD$lake =="Virginia", "7/09/2022",
                            ifelse(metaD$time_point =="3" | metaD$lake =="Virginia", "7/21/2022",
                            ifelse(metaD$time_point =="4" | metaD$lake =="Virginia", "8/04/2022",
                            "8/22/2022"
                            )))))))))))))))))))))))))))))

# adjust formatting to be Date (from)
metaD$date<-as.Date(mdy(metaD$date))

metaD<- metaD %>%
  dplyr::select(sequencing_ID, sampleNames, sample_ID, read.sum, time_point, date, lake, latitude, longitude, basin, sample_type, organism, Number.of.Organism, phy_group)

   
# the metadata above is now up to date with info necessary for downstream analysis
# now merge in the environmental data 

################################## ##################################
# read in the environmental data
env.metad<-read.csv("data/full.library.env.metaD.csv")

# make a column for 'Samplenames' and use this as rownames
env.metad$sampleNames<-env.metad$sequencing_ID
env.metad$sampleNames <- gsub("^.{0,3}", "", env.metad$sampleNames)
rownames(env.metad)<- env.metad$sampleNames

# remove the rows not found in the meta-meta data -- these are samples lost in dada2/controls
env.metad<-env.metad[(env.metad$sampleNames %in% metaD$sampleNames),]

env.metad<- env.metad %>%
  dplyr::select(sample_ID, sequencing_ID, sampleNames, sample_type, organism, Number.of.Organism, time_point, lake, 
                elevation_m, temp_C, chla_ug.L, pH, DO_perc, spc, cond, DOC, TDN, TDP)

################################## ##################################
# no merge, subset to make it easier to merge...
env.metad.reduce<-env.metad %>%
  dplyr::select(sampleNames, elevation_m, temp_C, chla_ug.L, pH, DO_perc, spc, cond, DOC, TDN, TDP)


# merge the two = final metadata, add in rownames for phyloseq merge
MetaD.SQ.Env<-merge(metaD, env.metad.reduce, by="sampleNames", all.x=TRUE)
rownames(MetaD.SQ.Env)<-MetaD.SQ.Env$sampleNames

# make new column for above or below treeline
MetaD.SQ.Env$elev.cat <- as.factor(ifelse(MetaD.SQ.Env$elevation_m < 2900, "Below", "Above"))

# are row names the same? if so, let's boogie
identical(rownames(sample_data(ps.prune.LOCAL)), rownames(MetaD.SQ.Env))
write.csv(MetaD.SQ.Env, "output/MetaD.SQ.Env.csv")

#updated metadata back into phyloseq
sample_data(ps.prune.LOCAL)<-MetaD.SQ.Env
PS.fin<- ps.prune.LOCAL

table(sample_data(PS.fin)$sample_type) #329 samples: 30 water, 299 zooplankton

# change to this order for later plot tests-- ultimately Bosmina and Ceriodaphnia are dropped nad large calanoids grouped into calanoids...
sample_data(PS.fin)$organism<-factor(sample_data(PS.fin)$organism, levels =c("Daphnia", "Holopedium", 
                                            "Calanoid", "Cyclopoid", "Water", "Bosmina", "Ceriodaphnia", 
                                            "Large Calanoid")) # back to factor

########### 
saveRDS(PS.fin, "output/reanalysis/PS.fin.RDS") # save phyloseq
```

Inspect the run: what is the # of reads/sample? A bit over inflated due to water being so deeply sequenced, but log-readsum gives us a good picture
```{r total reads by species boxplots raw data}

#### inspect plots
pdf(file="figures/reads.sample.pdf", height=4, width=7)
ggplot(MetaD.SQ.Env, aes(x=sample_type, y=read.sum, color=organism)) + geom_boxplot() + theme_bw()
dev.off() 

pdf(file="figures/log.reads.sample.pdf", height=4, width=12)
ggplot(MetaD.SQ.Env, aes(x=organism, y=log(read.sum), color=organism)) + geom_boxplot() + theme_bw()
dev.off() 

# read depth a bit low for zooplankton vs water, average depth is less than 2000 reads
ggplot(MetaD.SQ.Env[(MetaD.SQ.Env$sample_type=="Zooplankton"),], aes(x=sample_type, y=read.sum, color=organism)) +
  geom_boxplot() + 
  theme_bw()
```

Now look at read depth and rarefaction curves to determine how to standardize sampling depth
```{r Read depth plots and rarefaction curves}

# Histogram of sample read counts
hist.depth <- ggplot(MetaD.SQ.Env, aes(read.sum)) + 
  geom_histogram(color="gold4", fill= "gold2", binwidth = 10000) + 
  xlab("Read counts") +
  ggtitle("Sequencing Depth") + theme_classic()

hist.depth
dev.copy(pdf, "figures/hist.depth.pdf", height=7, width=9)
dev.off() 

# a different view
read_depth_violin <- ggplot(MetaD.SQ.Env, aes(x=1, y=read.sum)) +
  geom_violin() + ggtitle("Distribution of sample sequencing depth") + scale_y_log10() +
  xlab("Samples") + ylab("Read Depth") + geom_boxplot(width=0.1)

read_depth_violin
dev.copy(pdf, "figures/read_depth_violin.pdf", height=4, width=5)
dev.off()

# Rarefaction curve
count_table_filt <- as.data.frame(otu_table(PS.fin))
rarecurve((count_table_filt), step=50, cex=0.5, xlim=c(0,100000), ylab = "ASVs", label=FALSE)
#abline(v = 500, lty = "dotted", col="red", lwd=2)

dev.copy(pdf, "figures/rare.raw.pdf", height=4, width=8)
dev.off() 
```

Now we will subset the phyloseq objects and compare the influence of rarefaction for the phyloseq data.
- PS.fin is the final phyloseq object with all the data.  
- PS.500 is rarefied to 500 read-depth (with low abundance samples removed)
- PS.zoop is rarefied zooplankton data (500 read depth)
- PS.water is rarefied water only (500 read depth)
```{r exploring what low read count samples do and what to rarefy to}
###### subset the PS objects for needs...
# final and NOT rarefied data: 
PS.fin

# rarified to 500 reads: PS.rar, then reclassify and reduce some taxa and name "PS.500'
PS.rar = rarefy_even_depth(PS.fin, rngseed=111, sample.size=500, replace=F) 
table(sample_data(PS.rar)$sample_type) #288 samples: 28 water, 260 zooplankton


######################
# rarify, and hellinger
PS.500.hell.test = transform_sample_counts(PS.rar, function(x) x^0.5)

###### ###### ###### 
# Compare full data vs. rarefied data beta diversity
####### full data
ord.full <- ordinate(PS.fin, method = 'PCoA', distance = 'bray')
plot_ordination(PS.fin, ord.full, color = "sample_type")

# what samples have low reads
plot_ordination(PS.fin, ord.full, shape = "sample_type") +
  geom_point(aes(color = read.sum > 500))

Rare.culling<-plot_ordination(PS.fin, ord.full, color = "organism") +
  geom_point(aes(shape = read.sum > 500, size= read.sum > 500)) +
  scale_size_manual(values= c(6,2)) +
  scale_shape_manual(values=c(1, 16))+
  ggtitle("Samples culled by rarefaction") + theme_classic()

# plot the PCoA for non-rarified
Non.rar.NMDS<-plot_ordination(
  physeq = PS.fin,                                                   
  ordination = ord.full) +                                                
  geom_point(aes(color = organism), size = 2) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("ALL taxa-all data") +
  theme_classic() 

###### if rarefied
ord.rar <- ordinate(PS.rar, method = 'PCoA', distance = 'bray')
plot_ordination(PS.rar, ord.rar, color = "sample_type")

Rar.NMDS<- plot_ordination(
  physeq = PS.rar,                                                   
  ordination = ord.rar) +                                                
  geom_point(aes(color = organism), size = 2) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("ALL taxa- 500reads") +
  theme_classic() 

#### rarefied, but hellinger transformed
ord.hell <- ordinate(PS.500.hell.test, method = 'PCoA', distance = 'bray')

Hell.NMDS<- plot_ordination(
  physeq=PS.500.hell.test, 
  ordination = ord.hell) +
  geom_point(aes(color = organism), size = 2) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("ALL taxa - hellinger-500") +
  theme_classic() 


NMDS.tests<-plot_grid(Rare.culling + guides(color="none"), 
                      Non.rar.NMDS +  guides(color="none"), 
                      Rar.NMDS +  guides(color="none"), 
                      Hell.NMDS, ncol=4, rel_widths=c(4,4,4,5.5))
NMDS.tests
dev.copy(pdf, "figures/NMDS.tests.pdf", height=6, width=20)
dev.off()
#######

```

```{r final phyloseq objects}
################### final phyloseq object based on the above

# this PS.500 is used for the richness data -- non-rarefied
PS.500 = subset_samples(PS.rar, !organism=="Bosmina" & !organism=="Ceriodaphnia")
table(sample_data(PS.500)$organism) 

# classify the large calanoids as calanoids
sample_data(PS.500)$organism<-sample_data(PS.500)$organism %>% 
  dplyr::recode_factor(
    "Large Calanoid" = "Calanoid")

# check sample data
table(sample_data(PS.500)$organism) 

# set order of organims
sample_data(PS.500)$organism<-factor(sample_data(PS.500)$organism, levels =c("Daphnia", "Holopedium", 
                                            "Calanoid", "Cyclopoid", "Water")) # back to factor



##############################
######### hellinger transform the cleaned up data -- this is for beta diversity and NMDS etc
# rarify, and hellinger
PS.500.hell = transform_sample_counts(PS.500, function(x) x^0.5)

###### check the # of samples in each taxa
table(sample_data(PS.500.hell)$organism)
table(sample_data(PS.500.hell)$sample_type) 
#284 samples: 28 water, 256 zooplankton 
#4 samples removed for low-abundance

# sanity check
# rarefied, but hellinger transformed
ord.hell <- ordinate(PS.500.hell, method = 'PCoA', distance = 'bray')

NMDS.hell.rar.fin<- plot_ordination(
  physeq=PS.500.hell, 
  ordination = ord.hell) +
  geom_point(aes(color = organism), size = 2) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("ALL taxa - hellinger-500.pau") +
  theme_classic() 

NMDS.hell.rar.fin
dev.copy(pdf, "figures/NMDS.hell.rar.fin.pdf", height=6, width=7)
dev.off()
```

Here is the map of sampling locations
```{r map}
gps <- as.data.frame(cbind(unique(sample_data(PS.500.hell)$longitude), unique(sample_data(PS.500.hell)$latitude)))
rownames(gps) <- unique(sample_data(PS.500.hell)$lake)

map <- leaflet() %>%
  setView(lng = -119, lat = 38.25, zoom = 10) %>%
  addTiles() %>%
  fitBounds(-118.75, 37.42, -119, 38.06) %>%
  addMarkers(data = gps, lng = ~V1, lat = ~V2, popup = ~row.names(gps)) %>%
  addMarkers(
    lng = -118.7426, lat = 37.43153,
    label = "Eastern Brook (3,155 m)",
    labelOptions = labelOptions(noHide = T, direction = "left",
      style = list(
        "color" = "hotpink",
        "font-family" = "serif",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"))) %>%
  addMarkers(
    lng = -118.7444, lat = 37.43839,
    label = "Serene (3,124 m)",
    labelOptions = labelOptions(noHide = T, direction = "right",
      style = list(
        "color" = "green",
        "font-family" = "sanserif",
        "font-style" = "bold",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"))) %>%
  addMarkers(
    lng = -118.8574, lat = 37.59009,
    label = "Convict (2,315 m)",
    labelOptions = labelOptions(noHide = T, direction = "right",
      style = list(
        "color" = "orange",
        "font-family" = "serif",
        "font-style" = "bold",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"))) %>%
   addMarkers(
    lng = -119.2651, lat = 38.04698,
    label = "Virginia (2,983 m)",
    labelOptions = labelOptions(noHide = T, direction = "bottom",
      style = list(
        "color" = "gold",
        "font-family" = "serif",
        "font-style" = "bold",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"))) %>%
  addMarkers(
    lng = -119.2703, lat = 38.05095,
    label = "Blue (3,005 m)",
    labelOptions = labelOptions(noHide = T, direction = "right",
      style = list(
        "color" = "blue",
        "font-family" = "serif",
        "font-style" = "bold",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)"))) %>%
  addMarkers(
    lng = -119.2770, lat = 38.04806,
    label = "Cooney (3,116 m)",
    labelOptions = labelOptions(noHide = T, direction = "left",
      style = list(
        "color" = "purple",
        "font-family" = "serif",
        "font-style" = "bold",
        "box-shadow" = "3px 3px rgba(0,0,0,0.25)",
        "font-size" = "12px",
        "border-color" = "rgba(0,0,0,0.5)")))
map
```


Richness (observed and shannon) using PS.500 
- look at richness by sample type (water vs. zooplankton)
- look at how time and site effect richness
- how does richness differ for individual taxa compared to water (it is much lower...)
```{r alpha diversity plots - plot_richness}
#richness by organism (or sample type i.e, water) -- not transformed, but rarefied

richness.plot.samples <- plot_richness(PS.500, x="sample_type", measures=c("Observed", "Shannon"), color="sample_type") +
  labs(y= "Alpha Diversity Measure", x = "Sample Type") + theme_bw() + 
  geom_boxplot(aes(fill=sample_type), alpha=0.7) +
  geom_point(aes(color=sample_type), alpha=0.5) +
  scale_color_manual(values = c("lightsteelblue3", "navajowhite3"))+ guides(color="none") +
  scale_fill_manual(values = c("lightsteelblue2", "navajowhite2"))+ guides(fill="none") +
  theme(axis.text.x=element_text(size=9)) + theme(strip.text.x = element_text(size = 12))

richness.plot.samples$layers <- richness.plot.samples$layers[-1] # remove the first layer of points 
richness.plot.samples

#####
#richness by time_point
alpha.plot.time <- plot_richness(PS.500, x="time_point", measures=c("Shannon")) + 
  labs(y= "ASV Richness", x = "Sampling Time Point") + theme_bw() + 
  geom_point(color="gray15", alpha=0.5) +
  geom_boxplot(fill= "gray85", alpha=0.5) + 
  theme(strip.text.x = element_text(size = 12))

alpha.plot.time$layers <- alpha.plot.time$layers[-1] # remove the first layer of points 
alpha.plot.time

#Shannon diversity by Lake
alpha.plot.location <- plot_richness(PS.500, x="lake", measures=c("Shannon")) +
  labs(y= "Shannon Diversity", x = "Lake") + theme_bw() + 
  geom_point(color="gray15", alpha=0.5) +
  geom_boxplot(fill= "gray15", alpha=0.5) +
  ylim(0,6) + theme(strip.text.x = element_text(size = 12))

alpha.plot.location$layers <- alpha.plot.location$layers[-1] # remove the first layer of points 
alpha.plot.location

# group the above two plots
alpha.time.lake<-plot_grid(alpha.plot.time, alpha.plot.location)

#####
#richness by organisms (zoop species and water)
richness.plot.spp <- plot_richness(PS.500, x="organism", color="organism", measures=c("Observed", "Shannon")) +
  labs(y= "Alpha Diversity Measure", x = "Sample Type") + 
  geom_boxplot(aes(fill=organism), alpha=0.2) + guides(fill="none") +
  geom_point(alpha=0.5) +
  theme(axis.text.x=element_blank(), strip.text.x = element_text(size = 12)) + theme_bw()

richness.plot.spp$layers <- richness.plot.spp$layers[-1] # remove the first layer of points 
richness.plot.spp


######
# lets look at beta diversity and distance to centroid a bit: (zoop species and water)
# using the rarefied, hellinger data here since beta diversity metric
set.seed(10000)
bc_500.hell.bd <- phyloseq::distance(PS.500.hell, method = "bray") # rarefied and hellinger transformed
bc_beta.df <- as.data.frame(sample_data(PS.500.hell))

# one approach
bd<-betadisper(bc_500.hell.bd, bc_beta.df$organism, type="centroid")
plot(bd, ellipse = TRUE, hull = FALSE, label.cex = 0.4, main="Distance to Centroids")
dev.copy(pdf, "figures/bc_beta.df.pdf", height=6, width=6)
dev.off()

sqrt(dist(bd$centroids[,bd$eig>0])^2 - dist(bd$centroids[,bd$eig<0]^2))

# one approach
bds<-betadisper(bc_water.hell.bd, bc_water.df$organism, type="centroid")
plot(bds, ellipse = TRUE, hull = FALSE, label.cex = 0.4, main="Distance to Centroids")
dev.copy(pdf, "figures/bc_beta.df.pdf", height=6, width=6)
dev.off()

#
bd$group.distances # this is Average distance to centroid for each group
scores(bd, 1:2, display = "centroids") # this is the centroid for each group
bd.df <- data.frame(Distance_to_centroid = bd$distances, organism=bd$group) # dataframe

betadisp.org.plot<- ggplot(data=bd.df, aes(x=organism, y=Distance_to_centroid, color=organism)) +
  geom_boxplot(aes(fill=organism), alpha=0.2) + guides(fill="none") +
  geom_point(alpha=0.5) + 
  ylim(0,0.8) +
  guides(color=guide_legend(title=""))+ xlab("Sample Type")+ ylab("Distance to centroid") +
  ggtitle("Betadispersion") +
  theme(axis.text.x=element_blank()) + theme_bw() 

####
# combine all richness and betadispersion plots
alpha.plots<-plot_grid(alpha.time.lake, richness.plot.samples, richness.plot.spp + guides(color="none"),
                       betadisp.org.plot, ncol=4, rel_widths=c(11,8,11,8))
alpha.plots
dev.copy(pdf, "figures/alpha.plots.pdf", height=6, width=24)
dev.off()

```


```{r Bray curtis distance matrix}
# just zoops, from rarefied data, hellinger transformed
PS.zoops = subset_samples(PS.500.hell, sample_type=="Zooplankton")

# just water, from rarefied data, hellinger transformed
PS.water = subset_samples(PS.500.hell, sample_type=="Water")

###### 
# calculating bray curtis distances
# we use the hellinger transformation to avoid/deal with the arch effect
# we will use rarefied data (PS.500) for alpha diversity, but the transformed PS.500.hell for beta diversity

set.seed(10000)
bc <- phyloseq::distance(PS.fin, method = "bray") # raw full data

set.seed(10000)
bc_500 <- phyloseq::distance(PS.500, method = "bray") # rarefied data

set.seed(10000)
bc_500.hell <- phyloseq::distance(PS.500.hell, method = "bray") # rarefied and hellinger transformed


# both of the below are using rarefied and hellinger transformed data: PS.zoops and PS.water
set.seed(10000)
bc_zoops <- phyloseq::distance(PS.zoops, method = "bray")

set.seed(10000)
bc_water <- phyloseq::distance(PS.water, method = "bray")
```


```{r metaMDS, scores, and PERMANOVAS from BC distance matrices}
#################
# rarefied
set.seed(10000)
nmds.rarhell <- metaMDS(bc_500.hell,
          distance = "bray",
          k = 3,
          maxit = 999, 
          wascores = TRUE)

sp.scores.hell <- as.data.frame(scores(nmds.rarhell))
nmds.rarhell.df <- merge(sample_data(PS.500.hell), sp.scores.hell, by = 'row.names', all = TRUE)

broad.model.full <- adonis2(bc_500.hell ~ lake*time_point*organism, data = nmds.rarhell.df, permutations = 1000)
broad.model.full

env.model.full <- adonis2(bc_500.hell ~ latitude + elevation_m + temp_C + pH + cond + DOC + chla_ug.L + DO_perc, data = nmds.rarhell.df, permutations = 1000)
env.model.full

#test for beta dispersion
anova(betadisper(bc_500.hell, nmds.rarhell.df$lake)) #  significant (barely)
anova(betadisper(bc_500.hell, nmds.rarhell.df$time_point)) #  significant
anova(betadisper(bc_500.hell, nmds.rarhell.df$organism)) # very significant


##################
# zoops only
set.seed(10000)
nmds3 <- metaMDS(bc_zoops,
          distance = "bray",
          k = 3,
          maxit = 999, 
          wascores = TRUE)

species.scores3 <- as.data.frame(scores(nmds3))
nmds.merge3 <- merge(sample_data(PS.zoops), species.scores3, by = 'row.names', all = TRUE)

broad.model3 <- adonis2(bc_zoops ~ time_point*lake*organism, data = nmds.merge3, permutations = 1000)
broad.model3

env.model3 <- adonis2(bc_zoops ~ latitude + elevation_m + temp_C + pH + cond + DOC + chla_ug.L + DO_perc, data = nmds.merge3, permutations = 1000)
env.model3 # latitude is best 

anova(betadisper(bc_zoops, nmds.merge3$time_point)) # significant


#######################
# water only
set.seed(10000)
nmds4 <- metaMDS(bc_water,
          distance = "bray",
          k = 3,
          maxit = 999, 
          wascores = TRUE)

species.scores4 <- as.data.frame(scores(nmds4))
nmds.merge4 <- merge(sample_data(PS.water), species.scores4, by = 'row.names', all = TRUE)

broad.model4 <- adonis2(bc_water ~ time_point * lake, data = nmds.merge4, permutations = 1000)
broad.model4

env.model4 <- adonis2(bc_water ~ latitude + elevation_m + temp_C + pH + cond + DOC + chla_ug.L + DO_perc, data = nmds.merge4, permutations = 1000)
env.model4 #latitude, temp, chlorophyll

anova(betadisper(bc_water, nmds.merge4$time_point)) #not significant
anova(betadisper(bc_water, nmds.merge4$lake)) #not significant
```

MAY NEED TO DELETE OR REPURPOSE THIS SECTION
```{r Hellinger transformation, eval=FALSE}
# Hellinger (square root) transform the data!
PS.hell = transform_sample_counts(PS.fin, function(x) x^0.5)

set.seed(10000)
# Bray Curtis dissimilarity matrix
bc.hell <- phyloseq::distance(PS.hell, method = "bray")

# making NMDS ordination
set.seed(10000)
nmds.hell <- metaMDS(bc.hell,
          distance = "bray",
          k = 3,
          maxit = 999, 
          wascores = TRUE)
stress_result <- nmds.hell$stress #0.176
species.scores.hell <- as.data.frame(scores(nmds.hell))
nmds.merge.hell <- merge(sample_data(PS.hell), species.scores.hell, by = 'row.names', all = TRUE)


# test the effects of lake, time, and sample type (water, Cladocera, Copepoda) on the BC dissimilarities of the microbial communities
set.seed(10000)
hell.model <- adonis2(bc.hell ~ lake*time_point + phy_group, data = nmds.merge.hell, permutations = 1000)
hell.model

# test the effects of spatial and environmental variables on the BC dissimilarities of the microbial communities
adonis2(bc.hell ~ temp_C + pH + DO_perc + chla_ug.L + latitude + longitude + DOC + cond + elevation_m, data = nmds.merge.hell, permutations = 1000)
set.seed(10000)
hell.model.env <- adonis2(bc.hell ~ temp_C, data = nmds.merge.hell, permutations = 1000)
hell.model.env


install.packages('devtools')
library(devtools)
install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis") 
library(pairwiseAdonis)

# post hoc pairwise comparison to see what's driving the significance:
pairwise.adonis2(bc.hell ~ time_point*phy_group, data = nmds.merge.hell)
set.seed(10000)
pairwise.adonis2(bc.hell ~ phy_group, data = nmds.merge.hell)
pairwise.adonis2(bc.hell ~ lake, data = nmds.merge.hell)

# beta dispersion test
permutest(betadisper(bc.hell, nmds.merge.hell$lake), pairwise = T)
permutest(betadisper(bc.hell, nmds.merge.hell$phy_group), pairwise = T)
permutest(betadisper(bc.hell, nmds.merge.hell$temp_C), pairwise = T)
permutest(betadisper(bc.hell, nmds.merge.hell$time_point), pairwise = T)
# time is significant in the betadisper, sample_type and phy_group are not

#testing this anosim function out since time did not pass homogeneity test
#anosim(bc.hell, nmds.merge.hell$time_point, permutations = 1000)

```


```{r Env. effects and averages of ASV Richness and Diversity}

######### zoops AND water
# measure richness and shannon diversity
# get out of phyloseq
Shan.rich<-estimate_richness(PS.500, measures =c("Observed", "Shannon")) # using rarified transformed data
rownames(Shan.rich)<-sapply(str_remove_all(rownames(Shan.rich),"X"),"[") # will remove the X added to the rownames
Shan.rich.df<-merge(data.frame(sample_data(PS.500)), Shan.rich, by = "row.names") # merge 


# bell curve/ normally distributed, so we can use anova
hist(Shan.rich.df$Shannon, main="Shannon diversity", xlab="", breaks=10)
shapiro.test(Shan.rich.df$Shannon)

# Testing effects on Shannon diversity index
Anova(lm(Shannon ~ lake, data = Shan.rich.df))
shan.anova <- aov(lm(Shannon ~ lake, data = Shan.rich.df))
tukey <- TukeyHSD(shan.anova) 
cld <- multcompLetters4(shan.anova, tukey)


# testing effects of environmental variables on Shannon diversity
a.div.mod <- lm(Shannon ~ temp_C + DOC + elevation_m + chla_ug.L + pH + DO_perc + cond, data = Shan.rich.df)
a.div.AIC <- MASS::stepAIC(a.div.mod, direction = "both")
summary(lm(Shannon ~ elevation_m + DO_perc + cond, data = Shan.rich.df))

ggscatter(Shan.rich.df, x = "cond", y = "Shannon",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson", cor.coef.coord = c(0.65,2.98),
          xlab = "Temperature", ylab = "Zooplankton Community Shannon Diversity")


# summaryBy for median, mean, and standard error of the Shannon diversity by lake
lake_div <- summaryBy(Shannon ~ lake, data = Shan.rich.df, FUN = c(median, mean, se))


# plot Shannon diversity average across lakes
bargraph.CI(lake, Shannon, data = Shan.rich.df,
                     xlab = "Sample Type", ylab = "Shannon Diversity", cex.lab = 1.5, x.leg = 2,
                     col = "black", angle = 45, cex.names = 1.25,
                     density = c(0,20), legend = TRUE)

```


```{r environmental correlations}

cor.test(sample_data(PS.500.hell)$latitude, sample_data(PS.500.hell)$chla_ug.L)

ggscatter(sample_data(PS.500.hell), x = "elevation_m", y = "DOC",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "elevation (m)", ylab = "DOC")

ggscatter(sample_data(PS.500.hell), x = "temp_C", y = "DOC",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "temperature", ylab = "DOC")

#significant
ggscatter(sample_data(PS.500.hell), x = "elevation_m", y = "chla_ug.L", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "elevation (m)", ylab = "chla")

# significant
ggscatter(sample_data(PS.500.hell), x = "elevation_m", y = "temp_C", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "elevation (m)", ylab = "Temperature (C)")

# significant
ggscatter(sample_data(PS.500.hell), x = "latitude", y = "temp_C",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "latitude", ylab = "Temperature")

# significant
ggscatter(sample_data(PS.500.hell), x = "latitude", y = "DOC", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "latitude", ylab = "DOC")

#significant
ggscatter(sample_data(PS.500.hell), x = "latitude", y = "chla_ug.L", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "latitude", ylab = "chla")

ggscatter(sample_data(PS.500.hell), x = "DOC", y = "chla_ug.L", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "DOC", ylab = "chla")

#significant
ggscatter(sample_data(PS.500.hell), x = "temp_C", y = "pH", color = "lake",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "temp_C", ylab = "pH")

ggscatter(sample_data(PS.500.hell), x = "latitude", y = "DO_perc",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "latitude", ylab = "DO")

ggscatter(sample_data(PS.500.hell), x = "latitude", y = "chla_ug.L",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "latitude", ylab = "Chl a")

ggscatter(sample_data(PS.500.hell), x = "latitude", y = "cond",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "latitude", ylab = "Conductivity")

ggscatter(sample_data(PS.500.hell), x = "latitude", y = "pH",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "latitude", ylab = "pH")
```


```{r final envfit}
# get env. data
env.hell<- nmds.rarhell.df %>%
  dplyr::select(latitude, longitude, elevation_m, temp_C, chla_ug.L, pH, DO_perc, spc, cond, DOC, TDN, TDP)

# run the envfit function
fit.env.hell <- envfit(nmds.rarhell, env.hell, na.rm = TRUE)

scrs <- as.data.frame(scores(nmds.rarhell, display = "sites"))

# make the plot
spp.scrs <- as.data.frame(scores(fit.env.hell, display = "vectors"))
spp.scrs.test <- cbind(spp.scrs, r=fit.env.hell$vectors[2], pvals=fit.env.hell$vectors[4], Envfact = rownames(spp.scrs))
names(spp.scrs.test)[names(spp.scrs.test) == 'r'] <- 'r2' # rename bc vector has naming issues

# subset to keep only those Envfactors that are significant 
spp.scrs.signif<-spp.scrs.test[spp.scrs.test$pvals< 0.05, ]

# env fit with NMDS
NMDS.env <- ggplot(nmds.rarhell.df) +
  geom_point(nmds.rarhell.df, mapping = aes(x = NMDS1, y = NMDS2, colour = organism)) +
  coord_fixed() + ## need aspect ratio of 1!
  geom_segment(data = spp.scrs.signif,
               aes(x = 0, xend = NMDS1, y = 0, yend = NMDS2),
               arrow = arrow(length = unit(0.25, "cm")), colour = "grey") +
  geom_text(data = spp.scrs.signif, aes(x = NMDS1, y = NMDS2, label = Envfact),
            size = 3)+
  theme_classic()

```


```{r plotting lake elevations to visualize spatial differences}

elevations <- MetaD.SQ.Env %>%
  arrange(elevation_m) %>%
  mutate(lake = factor(lake, levels=c("Eastern Brook", "Serene", "Convict", "Virginia", "Blue", "Cooney"))) %>%
  ggplot(aes(x=lake, y=elevation_m, color = lake)) +
  geom_point(aes(size = 1.3)) +
  theme_classic() +
  xlab("Lake") +
  ylab("Elevation (m)") +
  ylim(0,3500) +
  theme(axis.text.x=element_text(size=20)) + theme(axis.title.x=element_text(size=20)) + 
  theme(axis.text.y=element_text(size=20)) + theme(axis.title.y=element_text(size=20)) +
  theme(axis.text.x = element_text(angle = 45, vjust=1, hjust=1)) +
  scale_color_manual(values = c("orange2", "violet", "darkorchid4", "darkgreen", "royalblue", "chartreuse2"))
elevations
dev.copy(pdf, "reanalysis/figures/elevations.pdf", height=5, width=5)
dev.off()
```


```{r mantel test, geographic distance}
# to calculate geographic distances between lakes
# geographic distance dissim matrix - adapted from weinstein et al 
gps <- as.data.frame(cbind(sample_data(PS.500.hell)$longitude, sample_data(PS.500.hell)$latitude))
spatial_dist <- distm(gps, fun = distGeo)
rownames(spatial_dist) <- sample_data(PS.500.hell)$lake
colnames(spatial_dist) <- sample_data(PS.500.hell)$lake

# converting distances to km 
spatial_dist_km <- spatial_dist/1000
dim(spatial_dist_km)
spatial_dist <- as.dist(spatial_dist)
spatial_dist_km <- as.dist(spatial_dist_km) # make distance matrix

# run the Mantel test
mantel <-vegan::mantel(spatial_dist_km, bc_500.hell)
# observation value in the output is the observed correlation (r) and shows sign (pos or neg corr.)

# plotting
dist.df <- as.data.frame(cbind(spatial_dist_km, bc_500.hell))

dist_plot <- ggplot(dist.df, aes(x=spatial_dist_km, y=bc_500.hell)) +
  geom_point(size=0.8,alpha=0.5) +
  geom_smooth(method = "glm", formula = y~x,
              method.args = list(family = gaussian(link = 'log')), lwd=0.7, se = T)+ 
  xlab("Geographic distance (km) between lakes") + 
  ylab("Bray-Curtis dissimilarities")
dist_plot
dev.copy(pdf, "reanalysis/figures/dist_plot.pdf", height=4, width=5)
dev.off() 



# Attempting to also run a Procrustes test
norm_distances <- decostand(spatial_dist_km, method = "normalize")
norm_dissimilarities <- decostand(bc_500.hell, method = "normalize")
std_distances <- decostand(norm_distances, method = "standardize")
std_dissimilarities <- decostand(norm_dissimilarities, method = "standardize")
# Perform Procrustes analysis
procrustes_results <- protest(std_distances, std_dissimilarities)
# Extract transformed coordinates
transformed_coordinates <- procrustes_results$points

```


```{r MDS plots by time point}
# Time 1

T1 <- subset_samples(PS.500.hell, time_point=="1")
ORD.T1 <- ordinate(T1, method='MDS', distance='bray')


MDS.ord.T1.lake<-plot_ordination(
  physeq = T1,                                                   
  ordination = ORD.T1) +                                                
  geom_point(aes(color = lake, shape=organism), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=lake)) +
  ggtitle("Time1") +
  scale_color_brewer(palette = "RdBu") +
  scale_fill_brewer(palette = "RdBu") +
  theme_classic()     
MDS.ord.T1.lake

MDS.ord.T1.org<-plot_ordination(
  physeq = T1,                                                   
  ordination = ORD.T1) +                                                
  geom_point(aes(color = organism, shape=lake), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("Time1") +
  theme_classic()     
MDS.ord.T1.org

# Time 2
T2<- subset_samples(PS.500.hell, time_point=="2")
ORD.T2 <- ordinate(T2, method='MDS', distance='bray')
MDS.ord.T2.lake<-plot_ordination(
  physeq = T2,                                                   
  ordination = ORD.T2) +                                                
  geom_point(aes(color = lake, shape=organism), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=lake)) +
  ggtitle("Time2") +
  scale_color_brewer(palette = "RdBu") +
  scale_fill_brewer(palette = "RdBu") +
  theme_classic()     
MDS.ord.T2.lake

MDS.ord.T2.org<-plot_ordination(
  physeq = T2,                                                   
  ordination = ORD.T2) +                                                
  geom_point(aes(color = organism, shape=lake), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("Time2") +
  theme_classic()     
MDS.ord.T2.org


# Time 3
T3<- subset_samples(PS.500.hell, time_point=="3")
ORD.T3 <- ordinate(T3, method='MDS', distance='bray')
MDS.ord.T3.lake<-plot_ordination(
  physeq = T3,                                                   
  ordination = ORD.T3) +                                                
  geom_point(aes(color = lake, shape=organism), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=lake)) +
  ggtitle("Time3") +
  scale_color_brewer(palette = "RdBu") +
  scale_fill_brewer(palette = "RdBu") +
  theme_classic()     
MDS.ord.T3.lake

MDS.ord.T3.org<-plot_ordination(
  physeq = T3,                                                   
  ordination = ORD.T3) +                                                
  geom_point(aes(color = organism, shape=lake), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("Time3") +
  theme_classic()     
MDS.ord.T3.org


# Time 4
T4<- subset_samples(PS.500.hell, time_point=="4")
ORD.T4 <- ordinate(T4, method='MDS', distance='bray')

MDS.ord.T4.lake<-plot_ordination(
  physeq = T4,                                                   
  ordination = ORD.T4) +                                                
  geom_point(aes(color = lake, shape=organism), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=lake)) +
  ggtitle("Time4") +
  scale_color_brewer(palette = "RdBu") +
  scale_fill_brewer(palette = "RdBu") +
  theme_classic()     
MDS.ord.T4.lake

MDS.ord.T4.org<-plot_ordination(
  physeq = T4,                                                   
  ordination = ORD.T4) +                                                
  geom_point(aes(color = organism, shape=lake), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("Time4") +
  theme_classic()     
MDS.ord.T4.org


# Time 5
T5 <- subset_samples(PS.500.hell, time_point=="5")
ORD.T5 <- ordinate(T5, method='MDS', distance='bray')
MDS.ord.T5.lake<-plot_ordination(
  physeq = T5,                                                   
  ordination = ORD.T5) +                                                
  geom_point(aes(color = lake, shape=organism), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=lake)) +
  ggtitle("Time5") +
  scale_color_brewer(palette = "RdBu") +
  scale_fill_brewer(palette = "RdBu") +
  theme_classic()     
MDS.ord.T5.lake

MDS.ord.T5.org<-plot_ordination(
  physeq = T5,                                                   
  ordination = ORD.T5) +                                                
  geom_point(aes(color = organism, shape=lake), size = 3) +    
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("Time5") +
  theme_classic()     
MDS.ord.T5.org


# now put them all together
combined.mds <- plot_grid(
          MDS.ord.T1.lake + theme(axis.title.y=element_text(size=8)) + ggtitle("Time 1: Lake Column"),
          MDS.ord.T1.org + theme(axis.title.y=element_text(size=8)) + ggtitle("Time 1: Taxa Column"),
          MDS.ord.T2.lake + theme(axis.title.y=element_text(size=8)),
          MDS.ord.T2.org + theme(axis.title.y=element_text(size=8)),
          MDS.ord.T3.lake + theme(axis.title.y=element_text(size=8)),
          MDS.ord.T3.org + theme(axis.title.y=element_text(size=8)),
          MDS.ord.T4.lake + theme(axis.title.y=element_text(size=8)),
          MDS.ord.T4.org + theme(axis.title.y=element_text(size=8)),
          MDS.ord.T5.lake + theme(axis.title.y=element_text(size=8)),
          MDS.ord.T5.org + theme(axis.title.y=element_text(size=8)),
          ncol=2)
combined.mds
dev.print(pdf, "reanalysis/figures/combined.mds.pdf", height=20, width=10)
dev.off()
 
```


```{r General NMDS (time, lake, basin, etc.)}

nmds.time <- ggplot(nmds.rarhell.df, aes(x=NMDS1, y=NMDS2, color = time_point)) +
  geom_point(aes(color = time_point), size = 2) + 
  stat_ellipse(level=0.9, lwd = 1, linetype = 2, aes(color=time_point)) +
  scale_color_manual(values = c("dodgerblue","coral", "purple2", "turquoise", "orange2")) + 
  ggtitle("by Time, pooled") +
  theme_bw()
 
nmds.lake <- ggplot(nmds.rarhell.df, aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = lake), size = 2) + 
  stat_ellipse(level=0.9, lwd = 1, linetype = 2, aes(color=lake)) +
  scale_color_brewer(palette = "RdBu") +
  scale_fill_brewer(palette = "RdBu") +
  ggtitle("by Lake, pooled") +
  theme_bw()

nmds.basin <- ggplot(nmds.rarhell.df, aes(x=NMDS1, y=NMDS2))+
  geom_point(aes(color = basin), size = 2) + 
  stat_ellipse(level=0.9, linetype = 2, aes(color=basin)) +
  ggtitle("by Basin, pooled") +
  theme_bw()

nmds.elevation <- ggplot(nmds.rarhell.df, aes(x=NMDS1, y=NMDS2, color = elev.cat)) +
  geom_point(aes(color = elev.cat), size = 2) + 
  stat_ellipse(level=0.9, linetype = 2, aes(color=elev.cat)) +
  scale_color_manual(values=c("orchid", "goldenrod2"))+
  ggtitle("by Elevation, pooled") +
  theme_bw()

nmds.organism <- ggplot(nmds.rarhell.df, aes(x=NMDS1, y=NMDS2, color = organism)) +
  geom_point(aes(color = organism), size = 2) + 
  stat_ellipse(level=0.9, linetype = 2, aes(color=organism)) +
  ggtitle("by Taxa, pooled") +
  theme_bw()

nmds.phy_group <- ggplot(nmds.rarhell.df, aes(x=NMDS1, y=NMDS2, color = phy_group)) +
  geom_point(aes(color = phy_group), size = 2) + 
  stat_ellipse(level=0.9, lwd = 1, linetype = 2, aes(color=phy_group)) +
  theme_bw() +
  scale_color_manual(values = c("orange", "firebrick", "royalblue3")) +
  ggtitle("by Phygroup, pooled")

phy_group_plots_combined <- plot_grid(
  nmds.time, nmds.lake, nmds.elevation,
  nmds.basin, nmds.organism, nmds.phy_group,  nrow=2, ncol=3, labels = c("A","B", "C", "D", "E", "F", ""))

phy_group_plots_combined
dev.print(pdf, "reanalysis/figures/NMDS.groupeffects.pdf", height=12, width=18)
dev.off()

```


```{r NMDS plots broken down by each lake and time point}
centroid <- nmds.rarhell.df %>% 
  group_by(time_point) %>%
  summarize(NMDS1 = mean(NMDS1), NMDS2 = mean(NMDS2))
str(nmds.rarhell.df)



nmds.convict <- ggplot(nmds.rarhell.df[nmds.rarhell.df$lake == "Convict", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point, shape = phy_group), size = 2.5) +
  stat_ellipse(level=0.9, lwd = 0.8, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("blue","red", "purple2", "turquoise", "orange2")) +
  theme_bw() + ggtitle("Convict") +
  theme(axis.text.x=element_text(size=13)) + theme(axis.title.x=element_text(size=11)) +
  theme(axis.text.y=element_text(size=13)) + theme(axis.title.y=element_text(size=11)) +
  theme(plot.title = element_text(size = 20)) +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11)) #change legend text font size
nmds.convict
dev.print(pdf, "reanalysis/figures/nmds.convict.pdf", height=10, width=15)
dev.off

nmds.east.brook <- ggplot(nmds.rarhell.df[nmds.rarhell.df$lake == "Eastern Brook", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point, shape = phy_group), size = 2.5) +
  stat_ellipse(level=0.9, lwd = 0.8, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("blue","red", "purple2", "turquoise", "orange2")) +
  theme_bw() + ggtitle("Eastern Brook") +
  theme(axis.text.x=element_text(size=13)) + theme(axis.title.x=element_text(size=11)) +
  theme(axis.text.y=element_text(size=13)) + theme(axis.title.y=element_text(size=11)) +
  theme(plot.title = element_text(size = 20)) +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11)) #change legend text font size
nmds.east.brook
dev.print(pdf, "reanalysis/figures/nmds.east.brook.pdf", height=10, width=15)
dev.off

nmds.serene <- ggplot(nmds.rarhell.df[nmds.rarhell.df$lake == "Serene", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point, shape = phy_group), size = 2.5) +
  stat_ellipse(level=0.9, lwd = 0.8, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("blue","red", "purple2", "turquoise", "orange2")) +
  theme_bw() + ggtitle("Serene") +
  theme(axis.text.x=element_text(size=13)) + theme(axis.title.x=element_text(size=11)) +
  theme(axis.text.y=element_text(size=13)) + theme(axis.title.y=element_text(size=11)) +
  theme(plot.title = element_text(size = 20)) +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11)) #change legend text font size
nmds.serene
dev.print(pdf, "reanalysis/figures/nmds.serene.pdf", height=10, width=15)
dev.off

nmds.cooney <- ggplot(nmds.rarhell.df[nmds.rarhell.df$lake == "Cooney", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point, shape = phy_group), size = 2.5) +
  stat_ellipse(level=0.9, lwd = 0.8, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("blue","red", "purple2", "turquoise", "orange2")) +
  theme_bw() + ggtitle("Cooney") +
  theme(axis.text.x=element_text(size=13)) + theme(axis.title.x=element_text(size=11)) +
  theme(axis.text.y=element_text(size=13)) + theme(axis.title.y=element_text(size=11)) +
  theme(plot.title = element_text(size = 20)) +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11)) #change legend text font size
nmds.cooney
dev.print(pdf, "reanalysis/figures/nmds.cooney.pdf", height=10, width=15)
dev.off

nmds.blue <- ggplot(nmds.rarhell.df[nmds.rarhell.df$lake == "Blue", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point, shape = phy_group), size = 2.5) +
  stat_ellipse(level=0.9, lwd = 0.8, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("blue","red", "purple2", "turquoise", "orange2")) +
  theme_bw() + ggtitle("Blue") +
  theme(axis.text.x=element_text(size=13)) + theme(axis.title.x=element_text(size=11)) +
  theme(axis.text.y=element_text(size=13)) + theme(axis.title.y=element_text(size=11)) +
  theme(plot.title = element_text(size = 20)) +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11)) #change legend text font size
nmds.blue
dev.print(pdf, "reanalysis/figures/nmds.blue.pdf", height=10, width=15)
dev.off

nmds.virginia <- ggplot(nmds.rarhell.df[nmds.rarhell.df$lake == "Virginia", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point, shape = phy_group), size = 2.5) +
  stat_ellipse(level=0.9, lwd = 0.8, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("blue","red", "purple2", "turquoise", "orange2")) +
  theme_bw() + ggtitle("Virginia") +
  theme(axis.text.x=element_text(size=13)) + theme(axis.title.x=element_text(size=11)) +
  theme(axis.text.y=element_text(size=13)) + theme(axis.title.y=element_text(size=11)) +
  theme(plot.title = element_text(size = 20)) +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=13), #change legend title font size
        legend.text = element_text(size=11)) #change legend text font size
nmds.virginia
dev.print(pdf, "phyloseq analysis/figures/nmds.virginia.pdf", height=10, width=15)
dev.off

combined.nmds.lakes <- ggarrange(nmds.convict,
          nmds.east.brook,
          nmds.serene,
          nmds.cooney,
          nmds.blue,
          nmds.virginia,
          plot_layout(guides="collect"), labels=NULL, common.legend = TRUE, 
          legend="bottom", align="hv", 
          font.label = list(size=10, color="black", family=NULL, position="top", labels="auto"))
combined.nmds.lakes
dev.print(pdf, "reanalysis/figures/combined.nmds.lakes.pdf", height=10, width=15)
dev.off

nmds.cladocera.time <- ggplot(nmds.rarhell.df[nmds.rarhell.df$phy_group == "cladocera", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point), size = 4) +
  stat_ellipse(level=0.9, lwd = 1, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("orangered", "turquoise2", "navy", "violetred", "lightslateblue")) + 
  theme(axis.text.x= element_text(size = 15)) +
  theme(axis.text.y= element_text(size = 15)) +
  theme_bw() +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=20), #change legend title font size
        legend.text = element_text(size=18)) +#change legend text font size 
  ggtitle("Cladocera")
nmds.cladocera.time
dev.print(pdf, "reanalysis/figures/nmds.cladocera.time.pdf", height=10, width=15)
dev.off

nmds.copepoda.time <- ggplot(nmds.rarhell.df[nmds.rarhell.df$phy_group == "copepoda", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point), size = 4) +
  stat_ellipse(level=0.9, lwd= 1, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("orangered", "turquoise2", "navy", "violetred", "lightslateblue")) + 
  theme_bw() +
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=20), #change legend title font size
        legend.text = element_text(size=18)) + #change legend text font size
  ggtitle("Copepoda")
nmds.copepoda.time
dev.print(pdf, "phyloseq analysis/figures/nmds.copepoda.time.pdf", height=10, width=15)
dev.off

nmds.water.time <- ggplot(nmds.rarhell.df[nmds.rarhell.df$sample_type == "Water", ], aes(x=NMDS1, y=NMDS2)) +
  geom_point(aes(color = time_point), size = 3) +
  stat_ellipse(level=0.9, lwd =1, linetype = 2, aes(color = time_point)) +
  scale_color_manual(values = c("orangered", "turquoise2", "navy", "violetred", "lightslateblue")) + 
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=20), #change legend title font size
        legend.text = element_text(size=18)) + #change legend text font size
  theme_bw() +
  ggtitle("Water")
nmds.water.time
dev.print(pdf, "phyloseq analysis/figures/nmds.water.time.pdf", height=10, width=15)
dev.off
  

```


```{r capscale and dbRDA --> having a hard time figuring this out}
# Trying to find environmental variables that signiificantly explained variation in the 
# 16S community composition

# Spatial variables were calculated using trend-surface analysis
# calculated second-order orthogonal polynomial terms (latitude, longitude, latitude   longitude, 
# latitude2, longitude2) from UTM coordinates for  each lake using the poly function in R.
library(sp)
install.packages('rgdal')
library(rgdal)
gps <- as.data.frame(cbind(sample_data(PS.500)$longitude, sample_data(PS.500)$latitude))
coordinates(gps) <- c("V1", "V2")
# Define the coordinate reference system (CRS) for the latitude and longitude data
proj4string(gps) <- CRS("+proj=longlat +datum=WGS84")
# Convert latitude and longitude to UTM coordinates
utm_coords <- spTransform(gps, CRS("+proj=utm +zone=11 +datum=WGS84"))
# Extract the UTM coordinates from the transformed object
utm_easting <- utm_coords@coords[, 1]
utm_northing <- utm_coords@coords[, 2]
# Create a data frame for the UTM coordinates
coord_df <- data.frame(cbind(utm_easting, utm_northing))
# Calculate second-order orthogonal polynomial terms for latitude and longitude
poly_terms <- poly(coord_df$utm_easting, coord_df$utm_northing, degree = 2)
# Extract the polynomial terms from the matrix
coord_df$latitude_2 <- poly_terms[, 1]
coord_df$longitude_2 <- poly_terms[, 2]
coord_df$latitude_longitude <- poly_terms[, 3]
coord_df$latitude_squared <- poly_terms[, 4]
coord_df$longitude_squared <- poly_terms[, 5]


# CAPSCALE
set.seed(10000)
capscale_result <- capscale(bc.hell ~ cond + chla_ug.L + DO_perc + pH + temp_C + DOC + elevation_m, dfun = "distance", data = nmds.merge.hell, distance = "bray")
capscale_model <- capscale(bc.hell ~ 1, data = nmds.merge.hell, distance = "bray")
# fitting global model, then performing stepwise selection with ordiRstep
mod <- ordiR2step(capscale_model, scope = capscale_result)
Anova(mod) # no matter the order of the explanatory variables in the original formula, the output of this shows that temp. is most important
summary(mod)
plot(mod)

# most parsimonious dbRDA model
set.seed(10000)
capscale_model <- capscale(bc.hell ~ temp_C, data = nmds.merge.hell, dfun = "distance", distance = "bray")
capscale_model_an <- Anova(capscale_model)
plot(capscale_model)


mod_scores <- as.data.frame(scores(mod, choices = c(1,2), display = "sites"))
merge.dbrda <- merge(mod_scores, nmds.merge.hell, by = "row.names")



cap_scores <- as.data.frame(scores(capscale_model, display = "sites"))
nmds.merge.hell$CAP1 <- cap_scores$CAP1
nmds.merge.hell$MDS1 <- cap_scores$MDS1
CAPxMDS_data <- data.frame(CAP1 = cap_scores$CAP1, MDS1 = cap_scores$MDS1, time = sample_data(PS.hell)$time_point, lake = sample_data(PS.hell)$lake, temp = sample_data(PS.hell)$temp_C)
CAPxMDS <- ggplot(CAPxMDS_data, aes(x = CAP1, y = MDS1, color = lake)) +
  geom_point() + stat_ellipse() +
  #geom_smooth(aes(color=lake, fill=lake), method = "lm", alpha = .2) +
  xlab("CAP1 (Environment)") + ylab("MDS1 (Bray-Curtis)") +
  theme_bw()
CAPxMDS

ordiplot(CAPxMDS_data[,c(1:2)], type="n", main=substitute(paste("")), cex.main=1, display="sites", xlim=c(-0.25, 0.8), cex.lab=0.8, cex.axis=0.8)
abline(h = 0, lty = "dotted")
abline(v = 0, lty = "dotted")
points(capscale_model, "sites", cex=0.8, pch=16, col = "gray62")
ordihull(environm_plot, groups=nmds.merge.hell$lake, draw="polygon", alpha=80, col = unique(lake_colors))
legend("topright", legend=levels(lake.group), cex=1, pch=16, col=unique(lake_colors), pt.cex=1, bty="n")
par.new=T




#dbRDA
set.seed(10000) 
upr <- dbrda(bc.hell ~ temp_C + elevation_m + cond + chla_ug.L + DOC + DO_perc + pH, data = nmds.merge.hell)
lwr <- dbrda(bc.hell ~ 1, data = nmds.merge.hell, na.action = 'na.omit')
# fitting global model, then performing stepwise selection with ordiRstep
mod_selec <- ordiR2step(lwr, scope = upr)
anova(mod_selec)
anova(mod_selec, by = "axis", permu = 200)
summary(mod_selec)
RsquareAdj(mod_selec)
screeplot(mod_selec)


dbRDA <- dbrda(formula = bc.hell ~ DOC, data = nmds.merge3, na.action = "na.omit")
plot(dbRDA)
```


```{r stacked bar plots}

########### plotting relative abundance, help from Megan Demmel, this is what I will use for analysis of relative abundances/ composition

#ASV_rel_abund <- transform_sample_counts(PS.fin, function(x) x/sum(x))
#pm <- psmelt(ASV_rel_abund)
pm <- psmelt(PS.500)
Abund <- aggregate(Abundance ~ sample_type + phy_group + lake + time_point + sequencing_ID + Class, data=pm, FUN=sum)
sam <- aggregate(Abundance ~ sample_type + phy_group + lake + time_point +sequencing_ID, data=pm, FUN=sum) # this should include everything Abund has except for Class

colnames(sam)[ncol(sam)] <- "tot.abund" # renaming the last column 'tot.abund' so that we can differentiate that from rel.abund

Abund.final <- dplyr::full_join(Abund, sam)
Abund.final$rel_abund <- Abund.final$Abundance/Abund.final$tot.abund
View(Abund.final)

Abund.final$Class[Abund.final$rel_abund <= 0.05] <- "Other (<5%)" # classes that have rel.abund less than 5% will be named as other

# plot by sample type
overall.comp.class <- ggplot(Abund.final, aes(x=phy_group, y=rel_abund, fill=Class, color=Class)) +
  geom_bar(aes(fill=Class, color=Class), stat="identity", position="fill") +
  xlab("Sample Type") +ylab("Relative Abundance (%)") + theme_bw() +
  theme(axis.text.x=element_text(size=11)) + theme(axis.title.x=element_text(size=13)) +
  theme(axis.text.y=element_text(size=11)) + theme(axis.title.y=element_text(size=13)) +
  scale_fill_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "tan", "navy", "royalblue3", "tan4", "lightsalmon", "gray54")) +
  scale_color_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "tan", "navy", "royalblue3", "tan4", "lightsalmon", "gray54"))
overall.comp.class
dev.copy(pdf, "reanalysis/figures/overall.comp.class.pdf", height=6, width=10)
dev.off()

# rel abund plot across time, lake, and sample type
Facet <- ggplot(Abund.final, aes(x = time_point, y = rel_abund, fill= Class, color = Class)) +
  geom_bar(width = 0.8, aes(fill = Class, color = Class), stat ="identity", position = "fill") +
  facet_grid(cols = vars(lake), rows = vars(phy_group)) + xlab("Sampling Time Point") +  ylab("Relative Abundance (%)") +
  theme(axis.text.x=element_text(size=11)) + theme(axis.title.x=element_text(size=13)) +
  theme(axis.text.y=element_text(size=11)) + theme(axis.title.y=element_text(size=13)) +
  theme(strip.text.x = element_text(size = 13)) + theme(strip.text.y = element_text(size = 13)) +
  scale_fill_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "tan", "navy", "royalblue3", "tan4", "lightsalmon", "gray54")) +
  scale_color_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "tan", "navy", "royalblue3", "tan4", "lightsalmon", "gray54"))
Facet
dev.copy(pdf, "reanalysis/figures/Facet.pdf", height=6, width=12)
dev.off()

# Help from Dr. Jackrel

my_class <- PS.500 %>%
  tax_glom(taxrank = "Class") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.05) %>%                         # Filter out low abundance taxa
  arrange(Class)  
view(my_class)

my_order <- PS.500 %>%
  tax_glom(taxrank = "Order") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.05) %>%                         # Filter out low abundance taxa
  arrange(Order)  
view(my_order)

# plot
overall.comp.class <- ggplot(my_class, aes(x=phy_group, y=Abundance, fill=Class, color=Class)) +
  geom_bar(aes(fill=Class, color=Class), stat="identity", position="fill") +
  ylab("Relative Abundance (%)") + theme(axis.text.x=element_text(size=12)) +
  theme(axis.text.y=element_text(size=10)) +
  theme(strip.text.y = element_text(size = 15)) + theme_bw() +
  scale_fill_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "skyblue4", "navy", "royalblue3", "tan3", "salmon2", "gray54")) +
  scale_color_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "skyblue4", "navy", "royalblue3", "tan3", "salmon2", "gray54"))
overall.comp.class
dev.copy(pdf, "reanalysis/figures/overall.comp.class.pdf", height=6, width=10)
dev.off()


overall.comp.order <- ggplot(my_order, aes(x=phy_group, y=Abundance, fill=Order, color=Order)) +
  geom_bar(aes(fill=Order, color=Order), stat="identity", position="fill") +
  ylab("Relative Abundance (%)") + theme(axis.text.x=element_text(size=12)) +
  theme(axis.text.y=element_text(size=10)) +
  theme(strip.text.y = element_text(size = 15)) + theme_bw() +
  scale_fill_manual(values = c('turquoise','pink2', 'tan', 'orange', 'cornflowerblue', 
                'magenta', 'darkolivegreen4', 'indianred1', 'tan4', 'darkblue', 
                'mediumorchid1','firebrick4',  'yellowgreen', 'lightsalmon', 'tan3',
                "tan1",'darkgray', 'wheat4', '#DDAD4B', 'chartreuse', 
                'seagreen1', 'moccasin', 'mediumvioletred', 'seagreen','cadetblue1',
                "darkolivegreen1" ,"tan2" ,   "tomato3" , "#7CE3D8","gainsboro", "gold2", "chocolate", "darkgreen", "dodgerblue", "burlywood3", 'hotpink', 'yellow3', 'blue', 'pink', 'brown3')) +
  scale_color_manual(values = c('turquoise','pink2', 'tan', 'orange', 'cornflowerblue', 
                'magenta', 'darkolivegreen4', 'indianred1', 'tan4', 'darkblue', 
                'mediumorchid1','firebrick4',  'yellowgreen', 'lightsalmon', 'tan3',
                "tan1",'darkgray', 'wheat4', '#DDAD4B', 'chartreuse', 
                'seagreen1', 'moccasin', 'mediumvioletred', 'seagreen','cadetblue1',
                "darkolivegreen1" ,"tan2" ,   "tomato3" , "#7CE3D8","gainsboro", "gold2", "chocolate", "darkgreen", "dodgerblue", "burlywood3", 'hotpink', 'yellow3', 'blue', 'pink', 'brown3'))
overall.comp.order
dev.copy(pdf, "reanalysis/figures/overall.comp.order.pdf", height=6, width=10)
dev.off()


Facet.class <- ggplot(my_class, aes(x = time_point, y = Abundance, fill= Class, color = Class)) +
  geom_bar(width = 0.8, aes(fill = Class, color = Class), stat ="identity", position = "fill") +
  facet_grid(cols = vars(lake), rows = vars(phy_group)) + xlab("Sampling Time Point") +  ylab("Relative Abundance (%)") +
  theme_bw() + theme(axis.text.x=element_text(size=12)) +
  theme(axis.text.y=element_text(size=10)) +
  theme(strip.text.y = element_text(size = 15)) +
  scale_fill_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "skyblue4", "navy", "tan3", "royalblue3", "black", "antiquewhite2", "chartreuse3", "chocolate", "orchid", "lightpink")) +
  scale_color_manual(values = c("red4", "orangered", "gold", "lightgreen", "turquoise3", "seagreen4", "lightskyblue", "skyblue4", "navy", "tan3", "royalblue3", "black", "antiquewhite2", "chartreuse3", "chocolate", "orchid", "lightpink"))
Facet.class
dev.copy(pdf, "reanalysis/figures/Facet.class.pdf", height=6, width=10)
dev.off()

```


```{r top average taxa compositions for each sample type}
# missing this data, can't compute
dominant_taxa <- summaryBy(rel_abund ~ Class*sample_type, data = Abund.final, FUN = mean)

```


```{r testing effects of env. variables on rel. abund. and richness of dominant bacteria}

# using Abund.final from the last code chunk
# testing the effects of time, lake, and sample type on the 5 most dominant classes of bacteria

# GAMMAPROTEOBACTERIA
gammaproteobacteria <- subset(Abund.final, Class == "Gammaproteobacteria")
gammaproteobacteria <- summaryBy(rel_abund ~ sequence_ID, data = gammaproteobacteria, FUN = sum)
gammaproteobacteria <- data.frame(gammaproteobacteria, row.names = gammaproteobacteria[,1])
gammaproteobacteria <- merge(gammaproteobacteria, sample_data(PS.fin), by = "row.names")
Anova(lm(rel_abund.sum ~ time_point + phy_group, data = gammaproteobacteria)) # time and phy

# testing environmental effects on gamaproteobacteria abundances
FitAll.gamma <- lm(rel_abund.sum ~ temp_C + DO_perc + pH + DOC + elevation_m + cond + chla_ug.L, data = gammaproteobacteria)
step(FitAll.gamma, direction = "both")
summary(lm(rel_abund.sum ~ DOC + elevation_m + cond + chla_ug.L, data = gammaproteobacteria)) # chlorophyll is significant but relies on other variables
gamma_chla <- ggplot(gammaproteobacteria, aes(x=chla_ug.L, y=rel_abund.sum)) +
  geom_point() +
  geom_smooth() + xlab("Chlorophyll a (ug/L)") + ylab("Gammaproteobacteria Abundance") +
  theme(axis.text.x=element_text(size=14)) + theme(axis.title.x=element_text(size=16)) +
  theme(axis.text.y=element_text(size=14)) + theme(axis.title.y=element_text(size=16))
gamma_chla
dev.copy(pdf, "phyloseq analysis/figures/gamma_chla.pdf", height=6, width=10)
dev.off()

gamma_chla <- ggscatter(gammaproteobacteria, x = "chla_ug.L", y = "rel_abund.sum",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list( method = "pearson"),cor.coef.coord = c(0.0,0.98),
          xlab = ("Chlorophyll a (ug/L)"), ylab = ("Gammaproteobacteria Abundance"))
gamma_chla 
dev.copy(pdf, "phyloseq analysis/figures/gamma_chla.pdf", height=6, width=10)
dev.off()


# ALPHAPROTEOBACTERIA
alphaproteobacteria <- subset(Abund.final, Class == "Alphaproteobacteria")
alphaproteobacteria <- summaryBy(rel_abund ~ sequence_ID, data = alphaproteobacteria, FUN = sum)
alphaproteobacteria <- data.frame(alphaproteobacteria, row.names = alphaproteobacteria[,1])
alphaproteobacteria <- merge(alphaproteobacteria, sample_data(PS.fin), by = "row.names")
Anova(lm(rel_abund.sum ~ lake + time_point + phy_group, data = alphaproteobacteria)) # time
TukeyHSD(aov(lm(rel_abund.sum ~ lake + time_point + phy_group, data = alphaproteobacteria)))
FitAll.alpha <- lm(rel_abund.sum ~ temp_C + DO_perc + pH + DOC + elevation_m + cond + chla_ug.L, data = alphaproteobacteria)
step(FitAll.alpha, direction = "both") # chla
summary(lm(rel_abund.sum ~ chla_ug.L, data = alphaproteobacteria)) #nope, chl a is not actually significant on its own


#ACTINOBACTERIA
actinobacteria <- subset(Abund.final, Class == "Actinobacteria")
actinobacteria <- summaryBy(rel_abund ~ sequence_ID, data = actinobacteria, FUN = sum)
actinobacteria <- data.frame(actinobacteria, row.names = actinobacteria[,1])
actinobacteria <- merge(actinobacteria, sample_data(PS.fin), by = "row.names")
Anova(lm(rel_abund.sum ~ lake*time_point*phy_group, data = actinobacteria)) # time
act.aov <- aov(lm(rel_abund.sum ~ lake + time_point + phy_group, data = actinobacteria))
TukeyHSD(act.aov)
FitAll.actino <- lm(rel_abund.sum ~ temp_C + DO_perc + pH + DOC + elevation_m + cond + chla_ug.L, data = actinobacteria)
step(FitAll.actino, direction = "both") # temp, DOm pH, cond, chl a
summary(lm(rel_abund.sum ~ temp_C + DO_perc + pH + DOC + cond + 
    chla_ug.L, data = actinobacteria)) # temp, cond, chl a

actino_DO <- ggplot(actinobacteria, aes(x=temp_C, y=rel_abund.sum)) +
  geom_point() +
  geom_smooth() + xlab("Dissolved Oxygen (%)") + ylab("Actinobacteria Abundance") +
  theme(axis.text.x=element_text(size=14)) + theme(axis.title.x=element_text(size=16)) +
  theme(axis.text.y=element_text(size=14)) + theme(axis.title.y=element_text(size=16))
actino_DO 
dev.copy(pdf, "phyloseq analysis/figures/actino_DO.pdf", height=6, width=10)
dev.off()

actino_DO <- ggscatter(actinobacteria, x = "temp_C", y = "rel_abund.sum",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list( method = "pearson"), cor.coef.coord = c(65,0.45),
          xlab = ("Dissolved Oxygen (%)"), ylab = ("Actinobacteria Abundance"))
actino_DO 
dev.copy(pdf, "phyloseq analysis/figures/actino_DO.pdf", height=6, width=10)
dev.off()

# CYANOBACTERIA
cyanobacteriia <- subset(Abund.final, Class == "Cyanobacteriia")
cyanobacteriia <- summaryBy(rel_abund ~ sequence_ID, data = cyanobacteriia, FUN = sum)
cyanobacteriia <- data.frame(cyanobacteriia, row.names = cyanobacteriia[,1])
cyanobacteriia <- merge(cyanobacteriia, sample_data(PS.fin), by = "row.names")
Anova(lm(rel_abund.sum ~ lake + time_point + phy_group, data = cyanobacteriia)) # time
FitAll.cyano <- lm(rel_abund.sum ~ temp_C + DO_perc + pH + DOC + elevation_m + cond + chla_ug.L, data = cyanobacteriia)
stepAIC(FitAll.cyano, direction = "both") # DOC

# plot
cyano_DOC <- ggplot(cyanobacteriia, aes(x=DOC, y=rel_abund.sum)) +
  geom_point() +
  geom_smooth() + xlab("DOC") + ylab("Cyanobacteria Abundance") +
  theme(axis.text.x=element_text(size=14)) + theme(axis.title.x=element_text(size=16)) +
  theme(axis.text.y=element_text(size=14)) + theme(axis.title.y=element_text(size=16))
cyano_DOC
dev.copy(pdf, "phyloseq analysis/figures/cyano_DOC.pdf", height=6, width=10)
dev.off()

```


```{r}
ASV.tab<-as.data.frame(otu_table(PS.500.hell))
ASV.samp<-sample_data(PS.500.hell)

ASV.data.all<-as.data.frame(cbind(ASV.samp,ASV.tab))

remotes::install_github("Russel88/MicEco")

# venns of all samples by time
All.venn.time<-ps_venn(
  PS.500.hell,
  "time_point",
  fraction = 0,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE,
  main="All times"
)

# venns of all phy
All.venn.phy<-ps_venn(
  PS.500.hell,
  "phy_group",
  fraction = 0,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE,
  main="All times"
)

# venn for all water
All.venn.water.time<-ps_venn(
  PS.water,
  "time_point",
  fraction = 0,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE,
  #fill= c("lightblue", "orange"),
  main="Water by time"
)


# venn for all daphnia
PS.daph = subset_samples(PS.zoops, organism=="Daphnia")
PS.daph.south = subset_samples(PS.zoops, basin=="South")
PS.daph.north = subset_samples(PS.zoops, basin=="North")

daph.S<-ps_venn(
  PS.daph.south,
  "lake",
  fraction = 0,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE,
  main="Daphnia by South-lake"
)

daph.N<-ps_venn(
  PS.daph.north,
  "lake",
  fraction = 0,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE,
  main="Daphnia by North-lake"
)

daph.basin<-ps_venn(
  PS.daph,
  "basin",
  fraction = 0,
  weight = FALSE,
  relative = TRUE,
  plot = TRUE,
  main="Daphnia by Basin"
)

g<-plot_grid(All.venn.time, All.venn.water.time, All.venn.phy,
             daph.N, daph.S, daph.basin, ncol=3)
g
dev.copy(pdf, "reanalysis/figures/Venns.pdf", height=8, width=12)
dev.off()


library("ggVennDiagram")
Venn.all<-ggVennDiagram(ASV.data.all, label_alpha=0) + 
  scale_fill_distiller(palette = "YlOrRd", direction = 1)+
  ggtitle("all data")

```

